---
title: "Homework_05"
author: "Joe Webb"
date: "2/16/2022"
output: html_document
---
Return to [Home Page](index.html)

For this lab, students explored the creation and alteration of **Matrices**, **Lists**, and **Data Frames**. The following work was completed in lab, and the output is posted as appropriate. 

## Question 1
Creating and messing around with a randomly generated matrix.

**Assign a random integer from 3 to 10 as`n_dims`.**
```{r setup, include=TRUE}
n_dims <- sample(3:10,1) 
print(n_dims)
```

**Creating a matrix from the pregenerated integer.**
```{r, include=TRUE}
# Create a vector from 1 to `n_dims^2`
vector <- c(1:n_dims^2) 

# Shuffle the vector's values
shuffled_vector <- sample(vector)

# Create a square matrix using the shuffled vector
matrix <- matrix(data=shuffled_vector, nrow=n_dims, ncol=n_dims, byrow=FALSE)
print(matrix)

# Transpose the matrix
trans_matrix <- t(matrix) 
print(trans_matrix) 
```
Transposing the matrix, flipped the `matrix` along its diagnol axis. 

**Extract statistical data from the matrix.**
```{r, include=TRUE}
# Name the matrix
rownames(trans_matrix) <- letters[1:n_dims]
colnames(trans_matrix) <- LETTERS[1:n_dims]
print(trans_matrix)

# Sum and mean the first and last row
# First Row
rowSums(trans_matrix)[1]
rowMeans(trans_matrix)[1]

# Last Row
rowSums(trans_matrix)[n_dims]
rowMeans(trans_matrix)[n_dims]
```

**Applying `eigen()` to the generated matrix.**
```{r, include=TRUE}
# Eigen
eigenmatrix <- eigen(trans_matrix)
print(eigenmatrix)

# Checking the typeof the retrieved Eeigen values.
print(typeof(eigenmatrix$values))
print(typeof(eigenmatrix$vectors))
```


## Question 2
**Creating a list to adapt and combine. **
```{r, include=TRUE}
# Creating a 4x4 matrix of random, uniform values
my_data <- runif(1:16)
my_matrix <- matrix(data=my_data, nrow=4, ncol=4, byrow=TRUE)
print(my_matrix)

# Creating a 100-Element Vector of True or False values
my_double <- runif(1:100)
head(my_double)

my_logical <- my_double < 0.5
head(my_logical)

# Creating a 26-element long vector or lower-case letters in random order
lowercase <- letters
my_letters <- sample(letters)
head(my_letters)

# Create a new list, which has the element[2,2] from the matrix, the second element of the logical vector, and the second element of the letters vector.
my_list <- list((my_matrix[2,2]), my_logical[2], my_letters[2])
print(my_list)

# Check the typeof each value in the list
print(my_list)
typeof(my_list[[1]])
typeof(my_list[[2]])
typeof(my_list[[3]])

# Take the list and create a vector with it.
new_vector <- unlist(my_list)
print(new_vector)

# What type is the vector?
print(typeof(new_vector))
```

## Question 3
**Creating a data set with letters and random uniform numbers.**
``` {r, include=TRUE}
# Creating variable 1: 26 random uniform numbers
my_unis <- runif(1:26)
head(my_unis)

# Creating variable 2: 26 capital letters
my_letters <- LETTERS
head(my_letters)

# Creating the data set using the two variables created abvoe.
my_df <- data.frame(my_unis, my_letters)

# Selecting and 4 Random Values in Column 1 with NA
my_df$my_unis[sample(nrow(my_df),4)]<-NA
print(my_df)

# Identify which rows received the NA
which(is.na(my_df), arr.ind=TRUE)

# Sort second variable alphabetically
my_df[order(my_df$my_letters), 2]
print(my_df)

# Calculate the column mean of the first variable
mean(my_df$my_unis, na.rm=TRUE)
```
