---
title: "Homework_07"
author: "Joe Webb"
date: "3/2/2022"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```
# Working with Dplyr
In this lab, students would open and manipulate the iris dataset. 

## Question 1
How many observations and variables are in this dataset?
```{r}
data(iris)
glimpse(iris)
```
There are 150 observations and 5 variables in this dataset. 

## Question 2
Create a new data frame `iris1` that contains only the species virginica and versicolor with sepal lengths longer than 6 cm and sepal widths longer than 2.5 cm. How many observations and variables are in the dataset?
```{r}
iris1 <- filter(iris, Species=="virginica"| Species=="versicolor", Sepal.Length > 6 & Sepal.Width > 2.5)

head(iris1)
```
`iris1` has 56 observations and 5 variables.

## Question 3
Now, create a `iris2` data frame from iris1 that contains only the columns for Species, Sepal.Length, and Sepal.Width. How many observations and variables are in the dataset?

```{r}
iris2 <- select(iris1, Species, Sepal.Length, Sepal.Width)
head(iris2)
```
`iris2` has 56 obserrvatins and 3 variables. 

## Question 4
Create an `iris3` data frame from `iris2` that orders the observations from largest to smallest sepal length. Show the first 6 rows of this dataset.
```{r}
iris3 <- arrange(iris2, by=desc(Sepal.Length))
head(iris3)
```

## Question 5
Create an `iris4` data frame from `iris3` that creates a column with a sepal area (length * width) value for each observation. How many observations and variables are in the dataset?
```{r}
Sepal_Area <- mutate(iris3, Sepal.Area=Sepal.Length*Sepal.Width)
iris4 <- select(Sepal_Area, 1:3, Sepal.Area)
head(iris4)
```
`iris4` has 56 observations and 4 variables. 

## Question 6
Create `iris5` that calculates the average sepal length, the average sepal width, and the sample size of the entire `iris4` data frame and print `iris5`.
```{r}
iris5 <- summarize(iris4, meanLength=mean(Sepal.Length), meanWidth=mean(Sepal.Width), sampleSize=n())
                   
print(iris5)
```

## Question 7
Finally, create `iris6` that calculates the average sepal length, the average sepal width, and the sample size *for each species* of in the `iris4` data frame and print `iris6`.
```{r}
irisSpecies <- group_by(iris4,Species)
iris6 <- summarize(irisSpecies, meanLength=mean(Sepal.Length), meanWidth=mean(Sepal.Width), sampleSize=n())
print(iris6)
```

## Question 8
In these exercises, you have successively modified different versions of the data frame `iris1` `iris1` `iris3` `iris4` `iris5` `iris6`. At each stage, the output data frame from one operation serves as the input fro the next. A more efficient way to do this is to use the pipe operator `%>%` from the `tidyr` package. See if you can rework all of your previous statements into an extended piping operation that uses iris as the input and generates `iris6` as the output.
```{r}
iris %>%
  filter(Species=="virginica"|Species=="versicolor", Sepal.Length>6 & Sepal.Width>2.5) %>%
  select(Species, Sepal.Length, Sepal.Width) %>%
  arrange(by=desc(Sepal.Length)) %>%
  group_by(Species) %>%
  summarize(meanLength=mean(Sepal.Length), meanWidth=mean(Sepal.Width), sampleSize=n())
```

## Quesion 9
Create a ‘longer’ data frame with three columns named: Species, Measure, Value.
```{r}
longiris <- iris %>% 
  pivot_longer(cols=Sepal.Length:Petal.Width, 
               names_to = "measure",
               values_to = "value")
               
head(longiris)
```