---
title: "Homework_06"
author: "Joe Webb"
date: "2/23/2022"
output: html_document
---
Return to [Homepage](index.html)

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(ggplot2) # for graphics
library(MASS) # for maximum likelihood estimation
```

## Simulating and Fitting Data Distributions
In this lab, students compared a histogram of continuous data to the probability density functions for different statistical distributions.

The following data is retrieved from: **The Supplemental Table S4, that accompany:  Rotenberry, J.T., and P. Balasubramaniam (2020).  Estimating egg mass-body mass relationships in birds.** Only the adult mass of 265 Accipitriformes were used in this example.


``` {r, echo=TRUE}
# Read the imported data file, consisting of 264 observations of adult Accipitriformes and their respective body masses. 

z <- read.table("bird_data.csv",header=TRUE,sep=",")
str(z)
summary(z)
```


``` {r, echo=TRUE}
# plot a histogram of the data

p1 <- ggplot(data=z, aes(x=adult_mass, y=..density..)) +
  geom_histogram(color="grey60",fill="cornsilk",size=0.2, bins = 25) 
print(p1)
```

``` {r, echo=TRUE}
# add empirical data curve
p1 <-  p1 +  geom_density(linetype="dotted",size=0.75)
print(p1)
```

``` {r, echo=TRUE}
normPars <- fitdistr(z$adult_mass,"normal")
print(normPars)
str(normPars)
normPars$estimate["mean"]

```

``` {r, echo=TRUE}
# plot normal probability density

meanML <- normPars$estimate["mean"]
sdML <- normPars$estimate["sd"]

xval <- seq(0,max(z$adult_mass),len=length(z$adult_mass))

stat <- stat_function(aes(x = xval, y = ..y..), fun = dnorm, colour="red", n = length(z$adult_mass), args = list(mean = meanML, sd = sdML))

print(p1 + stat)

```

``` {r, echo=TRUE}
#plot exponential probaility density

expoPars <- fitdistr(z$adult_mass,"exponential")
rateML <- expoPars$estimate["rate"]

stat2 <- stat_function(aes(x = xval, y = ..y..), fun = dexp, colour="blue", n = length(z$adult_mass), args = list(rate=rateML))
p1 + stat + stat2
```

``` {r, echo=TRUE}
# plot uniform probability density

stat3 <- stat_function(aes(x = xval, y = ..y..), fun = dunif, colour="darkgreen", n = length(z$adult_mass), args = list(min=min(z$adult_mass), max=max(z$adult_mass)))
p1 + stat + stat2 + stat3
```

``` {r, echo=TRUE}
# plot gamma probability density

z$am10 <- z$adult_mass/175
gammaPars <- fitdistr(z$am10,"gamma")
shapeML <- gammaPars$estimate["shape"]
rateML <- gammaPars$estimate["rate"]

stat4 <- stat_function(aes(x = xval, y = ..y..), fun = dgamma, colour="brown", n = length(z$adult_mass), args = list(shape=shapeML, rate=rateML))
p1 + stat + stat2 + stat3 + stat4
```

``` {r, echo=TRUE}
# plot beta probability density

pSpecial <- ggplot(data=z, aes(x=adult_mass/(max(adult_mass + 0.1)), y=..density..)) +
  geom_histogram(color="grey60",fill="cornsilk",size=0.2) + 
  xlim(c(0,1)) +
  geom_density(size=0.75,linetype="dotted")

betaPars <- fitdistr(x=z$adult_mass/max(z$adult_mass + 0.1),start=list(shape1=1,shape2=2),"beta")
shape1ML <- betaPars$estimate["shape1"]
shape2ML <- betaPars$estimate["shape2"]

statSpecial <- stat_function(aes(x = xval, y = ..y..), fun = dbeta, colour="orchid", n = length(z$adult_mass), args = list(shape1=shape1ML,shape2=shape2ML))
pSpecial + statSpecial
```

## Find Best Fitting Distribution
After examine the data, it is believed that **exponential probability curve** best fits the data. 

## Simulating the Data Set
``` {r, echo=TRUE}
#generate simulation values

simvector <- rexp(264, 1/1372.7)
simulation <- data.frame(ID= 1:264,
                         simulated_adult_mass=simvector)

simdata <- ggplot(data=simulation, aes(x=simulated_adult_mass, y=..density..)) +
  geom_histogram(color="grey60",fill="cornsilk",size=0.2, bins = 25) 

simdata + stat2
```
``` {r, echo=TRUE}
# copy of exponential graph

p1 + stat2
```
