---
title: "Homework_04"
author: "Joe Webb"
date: "2/9/2022"
output: html_document
---
Return to [Home Page](index.html)

# Question 1
Suppose x = 1.1, a = 2.2, and b = 3.3. Assign each expression to the value of the variable z and print the value stored in z.

```{r}
x <- 1.1
a <- 2.2
b <- 3.3
```

### 1a
```{r}
q1a<-x^a^b
print(q1a)
```

### 1b
```{r}
q1b<-(x^a)^b
print(q1b)
```

### 1c
```{r}
q1c<-3*x^3+2*x^2+1
print(q1c)
```


# Question 2
Using the `rep` and `seq` functions, create the following vectors:

### 2a
(1,2,3,4,5,6,7,8,7,6,5,4,3,2,1)
```{r}
q2b<-c(seq(from = 1, to = 8, by = 1), seq(from = 7, to = 1))
print(q2b)
```

### 2b
(1,2,2,3,3,3,4,4,4,4,5,5,5,5,5)
```{r}
q2b<-c(1,rep(2, 2), rep(3,3), rep(4, 4), rep(5,5))
print(q2b)

```

### 2c
(5,4,4,3,3,3,2,2,2,2,1,1,1,1,1)
```{r}
q2c<-c(5, rep(4,2), rep(3,3), rep(2,4), rep(1,5))
print(q2c)

```

# Question 3
Create a vector of two random uniform numbers. In a spatial map, these can be interpreted as x and y coordinates that give the location of an individual (such as a marked forest tree in a plot that has been mapped). Using one of R’s inverse trigonometry functions (asin(), acos(), or atan()), convert these numbers into polar coordinates.

```{r}
# Generating two random values between 0 and 1
q3_1<-runif(1) #this will be our x-value
q3_2<-runif(1) #this will be out y-value

# Plotting two points
library(ggplot2)
qplot(x=q3_1,y=q3_2)

# Converting to polar points
theta<-(atan(q3_2/q3_1)^-1)
print(theta)

r<-(q3_1^2+q3_2^2)
print(r)

```

# Question 4
Create a vector `queue <- c("sheep", "fox", "owl", "ant")` where `queue` represents the animals that are lined up to enter Noah’s Ark, with the sheep at the front of the line. Using R expressions, update `queue` as:
```{r}
queue <- c("sheep", "fox", "owl", "ant")

```

### 4a
the serpent arrives and gets in line;
```{r}
q4a<-append(queue, "serpent", after = 5)
print(q4a)

```

### 4b
the sheep enters the ark;
```{r}
typeof(q4a)

q4b <- q4a[! q4a %in% "sheep"]

print(q4b)

```

### 4c
the donkey arrives and talks his way to the front of the line;
```{r}
q4c<-append(q4b, "donkey", after = 0)

print(q4c)

```

### 4d
the serpent gets impatient and leaves;
```{r}
q4d <- q4c[! q4c %in% "serpent"]
print(q4d)

```

### 4e
the owl gets bored and leaves;
```{r}
q4e <-q4d[! q4d %in% "owl"]
print(q4e)

```

### 4f
the aphid arrives and the ant invites him to cut in line.
```{r}
q4f<- append(q4e, "aphid", after = 2)
print(q4f)

```

### 4g
Finally, determine the position of the aphid in the line.

```{r}
which(q4f=="aphid")

```

# Question 5
Use R to create a vector of all of the integers from 1 to 100 that are not divisible by 2, 3, or 7.

```{r}
vector <- (1:100)
q5<-which(vector %%2!=0 & vector %%3!=0 & vector %%7!=0)

print(q5)
```
